\chapter{Using \sysname{}}

There are several use cases for \sysname{} depending on the type and
level of ambition of the client.

\section{Implementing standard environment functions}

Unfortunately, the \commonlisp{} standard does not contain a separate
chapter dedicated environment functions.%
\footnote{The chapter of the standard
  entitled ``Environment'' has to do with the computational environment
  of the \commonlisp{} system.}
So for example the function \texttt{symbol-function} is specified in
the chapter entitled ``Symbols'', whereas the function
\texttt{fdefinition} is specified in the chapter entitled ``Data and
Control Flow'', even though the two are very much related.

In this section, we describe how a \commonlisp{} implementation could
implement the standard environment functions using a \sysname{}
run-time environment as a first-class global environment.  To access
the \sysname{} run-time environment, we are going to assume it is
available in the host special variable \texttt{*e*}, but in
a real system, accessing the value of a special variable may require
access to the global environment, so some other mechanism might be
needed.  Similarly, we assume that some kind of client object is
available in the special variable \texttt{*c*}.

In many cases, the \commonlisp{} standard does not specify that an
error must be signaled in situations that are not well defined, and in
some cases, it specifies that an error must be signaled, but only in
safe code.  To simplify our examples, we always include code to signal
an error i such situations, but our code may not be very specific in
terms of what error condition is signaled.  When the standard requires
an argument to an operator to be a function name, we assume that the
client has a global type definition \texttt{function-name} for
function names.

In our examples, we use the package prefix \texttt{env:} to refer to
\sysname{} symbols.

\subsection{\texttt{fboundp}}

The standard says that \texttt{fboundp} returns a true value if and
only if the name is defined either as a function, a macro, or a
special operator in the global environment.  We use the \sysname{}
functions \texttt{fdefinition}, \texttt{macro-function}, and
\texttt{special-operator} to query the environment.  Each of these
functions returns \texttt{nil} if no definition is present, so if all
of them return \texttt{nil}, then the name is not fbound.  The
standard says that, in safe code, the implementation must check that
the name is a valid function name, and we include code to check that
this is the case.

\begin{verbatim}
(defun fboundp (name)
  (check-type name function-name)
  (not (and (null (env:fdefinition *c* *e* name)))
            (null (env:macro-function *c* *e* name))
            (null (env:special-operator *c* *e* name))))
\end{verbatim}

\subsection{\texttt{fdefinition} and \texttt{symbol-function}}

The standard leaves several decisions about the behavior of
\texttt{fdefinition} up to the implementation.  It is specified that a
function object is returned if the name is defined as a function.  If
the name is defined as a macro or a special operator, the standard
specified that a true value must be returned, but it leaves the nature
of the value up to the implementation.  In our example implementation,
we assume that the \commonlisp{} implementation calls the \sysname{}
function \texttt{(setf special-operator)} with a value that is
acceptable as a return value for the standard function
\texttt{fdefinition} when the name is that of a special operator.

\begin{verbatim}
(defun fdefinition (name)
  (check-type name function-name)
  (let ((function (env:fdefinition *c* *e* name)))
    (if (not (null function))
        function
        (let ((macro (env:macro-function *c* *e* name)))
          (if (not (null macro))
              `(:macro ,macro)
              (let ((special (env:special-operator *c* *e* name)))
                (if (not (null special))
                    special
                    (error 'undefined-function ...))))))))
\end{verbatim}

The standard function \texttt{symbol-function} is like
\texttt{fdefinition} except that the name is restricted to be a
symbol, which the implementation must verify in safe code.

\begin{verbatim}
(defun symbol-function (name)
  (check-type name symbol)
  (fdefinition name))
\end{verbatim}

\subsection{\texttt{(setf fdefinition)} and \texttt{(setf symbol-function)}}

In the ``Notes'' section of the \commonlisp{} standard, it is said
that ``It is an error to set the \texttt{fdefinition} of a function
name to a symbol, a list, or the value returned by
\texttt{fdefinition} on the name of a macro or special form.''
However, the ``Notes'' section is not normative, and there are no
exceptional situations listed that related to the value of the
\textit{new-definition} argument.  This is a case where the standard
does not specify some error situations that might seem reasonable.
Our example here assumes that \textit{new-definition} must be a
function, or else an error is signaled.

Also, in the ``Notes'' section, it is indicated that \texttt{(setf
  fdefinition)} can be used only when \textit{function-name} does not
represent a special operator, but no exceptional situation for this
case is indicated.  In our example, we signal an error for this case.

\begin{verbatim}
(defun (setf fdefinition) (new-definition function-name)
  (check-type function-name function-name)
  (check-type new-definition function)
  (if (env:special-operator *c* *e* function-name)
      (error ...)
      (setf (env:fdefinition *c* *e* function-name)
            new-definition)))
\end{verbatim}

\subsection{\texttt{macro-function}}

The \sysname{} function behaves the same way as the standard function.

\begin{verbatim}
(defun macro-function (symbol &optional environment)
  (env:macro-function
    *c*
    (if (null environment) *e* environment)
    symbol))
\end{verbatim}

\subsection{\texttt{special-operator-p}}

The \sysname{} function behaves the same way as the standard function.

\begin{verbatim}
(defun special-operator-p (symbol)
  ;; At least in safe mode, check that SYMBOL is a symbol.
  (env:special-operator-p *c* *e* symbol))
\end{verbatim}

\section{Compilation environment only}

An existing (perhaps relatively new) \commonlisp{} implementation that
has decided to take advantage of the permission to merge the run-time
environment and the evaluation environment, but that wants a separate
compilation environment can use \sysname{} as follows:

\begin{itemize}
\item It would define an environment class  (say
  \texttt{client:global-environment}) that has
  \texttt{clostrum:run-time-environment} as a superclass.
\item It would implement methods on the generic functions in
  \refSec{sec-run-time-protocol-functions} that trampoline to existing
  implementation-specific functions in the single global environment
  of the client.
\item It would create a constellation consisting of an instance of
  \texttt{client:global-environment} as the parent of an instance of
  \texttt{clostrum:compilation-environment}
\end{itemize}

\section{Compilation and evaluation environment}

An existing \commonlisp{} implementation with a traditional, single
global run-time environment that wants a separate evaluation
environment in order to avoid compile-time side effects to its
run-time environment can use \sysname{} as follows:

\begin{itemize}
\item It would define an environment class  (say
  \texttt{client:run-time-environment}) that has
  \texttt{clostrum:environment} as a superclass.
\item It would implement methods on the generic functions in
  \refSec{sec-run-time-protocol-functions} that trampoline to existing
  implementation-specific functions in the single global environment
  of the client.
\item It would create a class (say
  \texttt{client:evaluation-environment}) as a subclass of
  \texttt{clostrum:evaluation-environment-mixin} and
  \texttt{client:run-time-environment}.
\item It would create a constellation consisting of an instance of
  \texttt{client:run-time-environment}, an instance of
  \texttt{client:evaluation-environment},
  \texttt{clostrum:compilation-environment}
\end{itemize}
